Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-24 arxiv/diff_cov.tex
wip

As mentioned in Section \ref{sec_prelim}, the optimization objective \eqref{eq_mtl} is non-convex in general.
Therefore, we consider an arbitrary local minimum and provide generalization bounds that scale with key properties of task data such as sample sizes and covariate shifts.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates in hard parameter sharing. %We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.

We first provide a key technical tool that we derived using recent developments in random matrix theory \cite{??}.
To motivate our study, we consider a specialized setting where the output layer weights $A_1, A_2$ are both equal to one.
Without loss of generality, we consider task two.
For this setting, we can obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ by solving $B$ in equation \eqref{eq_mtl} as follows:
\begin{align*}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}
\end{align*}
Next, we derive the prediction loss of hard parameter sharing in expectation over the random noise, since the empirical quantity is close to the expectation similar to what we have shown in Claim \ref{claim_pred_err}:
\begin{align*}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	\bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2
	+ \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}.
\end{align*}

>>>>

When the covariates across tasks are different, we can no longer obtain an optimal solution for minimizing equation \eqref{eq_mtl} as in Claim \ref{lem_exp_opt}.
This is because the minimization problem is non-convex with respect to $A$ and $B$.
Therefore, we consider an arbitrary local minimum of equation \eqref{eq_mtl} and provide generalization bounds that scale with key properties of task data such as sample sizes and covariate shifts.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates in hard parameter sharing. %We derive deterministic conditions to compare the prediction loss of the local minimum to single-task learning.
We will develop the key technical results in this section, and describe their implications in the next section.

We first provide a key technical tool that we derived using recent developments in random matrix theory \cite{??}.
To motivate our study, we consider a specialized setting where the output layer weights $A_1, A_2$ are both equal to one.
Without loss of generality, we consider the prediction loss of hard parameter sharing for task two and the same analysis applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align*}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align*}
For the prediction loss, since the empirical loss is close to the expected loss similar to what we have shown in Claim \ref{claim_pred_err}, we consider the expected loss conditional on the covariates as follows:
\begin{align*}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	\bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2
	+ \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}.
\end{align*}

### 2020-09-24 arxiv/setup.tex
wip

\begin{fact}\label{lem_minv}
	Let $c$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c})$, cf. Theorem 2.4 in \citet{isotropic}.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c}$.
	\end{enumerate}
\end{fact}

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise component, we assume that all of its moments exist.
More precisely, there exists a fixed function $C(\cdot) : \mathbb{Z} \rightarrow \real^+$ such that for any $a = 1, 2, \dots, \infty$, we have that
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{j}^{(i)}}^a} \le C(a), \text{ for any } 1\le i\le t \text{ and } 1\le j\le n_i.
\end{align}

>>>>

\begin{fact}\label{lem_minv}
	Let $c$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c})$, cf. Theorem 2.4 in \citet{isotropic}.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c}$, cf. \citet[Theorem 2.10]{isotropic} and \citet[Lemma 3.12]{DY}.
	\end{enumerate}
\end{fact}

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise component, we assume that all of its moments exist.
More precisely, there exists a fixed function $C(\cdot) : \mathbb{Z} \rightarrow \real^+$ such that for any $a = 1, 2, \dots, \infty$, we have that
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{j}^{(i)}}^a} \le C(a), \text{ for any } 1\le i\le t \text{ and } 1\le j\le n_i.
\end{align}
Hence, for any value $\varphi > 4$, we get that Fact \ref{lem_minv} holds for $\varepsilon^{(i)}$, for all $i = 1, 2, \dots, t$.
Let $\e > 0$ denote \todo{?? to check}.
